---
title: "HW1"
author: "Joe Zhang"
---

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(mclust)
library(survival)
```

# Problem 5
```{r, message=FALSE, warning=FALSE}
data <- read.csv('cell.csv')
```

## a)
It looks like the customers are split evenly across the months and that the customers each month are pretty close to splitting evenly between churn and no churn. According to the crosstable result, we should use stratified sampling method.
```{r, message=FALSE, warning=FALSE}
data <- cell
table(data[,c('churn','billmonth')])
```

## b)
2476 people canceled.
The retention rate equals 0.02589308.
```{r, message=FALSE, warning=FALSE}
sum(data$churn)
r_hat = sum(data$churn,na.rm = TRUE)/sum(data$t2,na.rm = TRUE)
r_hat
```

## c)
We have to do stratetified sampling. As the retention rate is different across account types or line counts. And the distribution of the value in those two features are not uniform. If we just randomly drwa sample from the dataset, it may give a lot of bias.

## d)
The average monthly revenue is 55.7633.
E(CLV) = m(1+d)/(1+d-r) = 57.23051
```{r, message=FALSE, warning=FALSE}
mean.rm.na <- function(x){
return(mean(x,na.rm = TRUE))}
function()mean(x,na.rm = TRUE)
tmp = apply(data[,7:21], 2, mean.rm.na)
m = mean(tmp)
d = 0.01
mean(tmp)
m*(1+d)/(1+d-r_hat)
```

## e) 
There is a significant drop at the 23-25 month mark which may indicate that there is a 2 year contract in place.
```{r, message=FALSE, warning=FALSE}
fit = survfit(Surv(t2, churn) ~ 1, data=data)
summary(fit)
plot(fit)
```

## f)
The individual is more likely to churn.
```{r, message=FALSE, warning=FALSE}
fit = survfit(Surv(t2, churn) ~ ACCOUNT_TYPE, data=data)
summary(fit)
plot(fit, col=1:2)
legend("topright", paste(" ",c("Business","Individual")), col=1:2, lty=c(1,2))
```
## g)

The fewer lines a customer/business may have, the more likely they will churn. Thos with more lines tend to churn less especially at the 24 month time.
```{r, message=FALSE, warning=FALSE}
data$LINE_COUNT2 = data$LINE_COUNT
data$LINE_COUNT2[data$LINE_COUNT2>5]=5
fit = survfit(Surv(t2, churn) ~ LINE_COUNT2, data=data)
summary(fit)
#plot(fit)
plot(fit, col=1:6)
legend("topright", paste(" ",c(0,1,2,3,4,5)), col=1:6, lty=c(1,6))
```