---
title: "Exploratory Factor Analysis"
author: "Kristiyan Dimitrov"
date: "3/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

```{r}
# Libraries
library(psych)
library(stats)
# Import Data
music = read_csv('music.csv')
music = as.matrix(music)
head(music)
```

## a) How many factors?

Should use scree plot and Kaiser criterion.
To make a scree plot, need the eigen values of the covariance matrix.
Can get this with prcomp()

```{r}
# ?prcomp
fit = prcomp(music)
fit$sdev^2
plot(1:28,fit$sdev^2, type='b') # This is the scree plot; Taking the squares of the sdev's, which are the eigenvalues
abline(a=1, b=0)
```

Second way of getting scree plot.

```{r}
screeplot(prcomp(music))
abline(a=1, b=0)
```

The Kaiser criterion would suggest more factors, but the elbow in the scree plot suggests that 3 factors should be enough.

## b) Using 3 factors for PCA

The professor switches to using principal() so he can specify the number of factors

```{r}
?principal
fit2 = principal(music, nfactors=3, rotate = 'none')
fit2
```

Looking at the output below, it looks like I should drop: V28, V29, V31, V37, V44, V47, V51, V53, V54 due to significant cross-loading

```{r}
fit3 = principal(music, nfactors=3, rotate = 'varimax')
print(fit3$loadings, cutoff=0.4, digits=3,
sort=TRUE)
```

Dropping some questions
```{r}
music = music[, !colnames(music) %in% c('V28',' V29', 'V31', 'V44', 'V47', 'V51', 'V53', 'V54')]
head(music)
```

Based on the new output, I'll drop: 'V29', 'V32','V50','V40','V45','V52','time1'

```{r}
fit4 = principal(music, nfactors=3, rotate = 'varimax')
print(fit4$loadings, cutoff=0.4, digits=3,
sort=TRUE)
```

Dropping mode questions

```{r}
music = music[, !colnames(music) %in% c('V29', 'V32','V50','V40','V45','V52','time1')]
head(music)
```

Doing PCA/EFA again

```{r}
fit5 = principal(music, nfactors=3, rotate = 'varimax')
print(fit5$loadings, cutoff=0.4, digits=3,
sort=TRUE)
```
## c) Naming the factors
Music is beneficial for me: Factor 1: 35,36,38,39,41,42,48 / Malthouse got 38, 39, 41, +37
I brand myself with Music: Factor 2: 30,33,34,43,46,49 / Malthouse got 33,34,43,46,49 (all but 30)
Emotional: Factor 3: 49 / Malthouse got 49, + 44, 45

## d) We haven't talked about alphas

## e)

```{r}
fit5 = principal(music, nfactors=3, rotate = 'varimax')
fit5$values
print(fit5$loadings, cutoff=0.4, digits=3,
sort=TRUE)
```

## f) correlated factors

```{r}
fit6 = principal(music, nfactors=3, rotate = 'promax')
fit6
```

The output gives us the correlation
     RC1  RC2  RC3
RC1 1.00 0.59 0.58
RC2 0.59 1.00 0.49
RC3 0.58 0.49 1.00

##g) Regressing on Time

```{r}
fit7 = prcomp(music, scale = T)
musicOrig = read_csv('music.csv')
musicOrig = as.matrix(musicOrig)
musicOrig
PCdata = data.frame('PC1' = fit7$x[,1], 'PC2'=fit7$x[,2], 'PC3'=fit7$x[,3], 'time' = musicOrig[,'time1'])
PCdata
fit8 = lm(time ~ PC1+PC2+PC3, PCdata)
summary(fit8)
```

