---
title: "Migration HW 6"
author: Kristiyan Dimitrov, Srividya Ganapathi, Shreyashi Ganguly, Greesham Simon, Joe Zhang
date: "3/08/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

```{r, message=FALSE, warning=FALSE}
gene <- read.csv("gene.csv")
```

## Part (a)

Run PCA on the data. What fraction of variation is accounted for by the first two components?

```{r, message=FALSE, warning=FALSE}
#Checking to see if all fields are measured on comparable scales
mean <- apply(gene[,-1001],2,mean)
hist(mean)
sd <- apply(gene[,-1001],2,sd)
hist(sd)
```

All the X variables are measurements of comparable units. Also, almost all the fields have mean 0 and standard deviation 1, hence not standardizing.

```{r, message=FALSE, warning=FALSE}
fit <- prcomp(gene[,-1001], scale=F)
plot(fit) #scree plot
summary(fit)
```

Proportion of variation accounted for by first two principal components are 15.9%

## Part (b)
Generate a scatterplot of the second component against the first. En-code sick using different colors and/or plotting symbols. Submit your scatterplot and briefly discuss what the plot reveals.

```{r}
plot(fit$x[,1:2], 
     pch=ifelse(gene$sick==1,15,19),
     col=ifelse(gene$sick==1,"red","green")) 
```

The plot reveals that the first two principal components can adequately capture the distinction between sick and healthy tissue samples


# Problem 2

```{r}
X=matrix(c(1:4, 1,4,9,16), nrow=4)
print("X:")
print(X)

XtX = t(X) %*% X
print("XtX:")
print(XtX)

XXt = X %*% t(X)
print("XXt:")
print(XXt)

fit_XtX = eigen(XtX)
print(fit_XtX, digits=2)

fit_XXt = eigen(XXt)
print(fit_XXt, digits=2)

```

$$ X -> 4 \times 2$$
$$ => X'X -> 2 \times 2$$
$$ => XX' -> 4\times4$$

The first two eigen values of $XX'$ are the same as that of $X'X$
$XX'$ has two additional eigen values (owing to it's 4x4 structure), but they are approximately 0

# Problem 3

```{r}
X = matrix(c(1,3,5,0,1,   2,4,4,2,3,   3,5,3,4,5), nrow=5)
print("X:")
print(X)
```

### Part (a)

```{r 5a, message=FALSE, warning=FALSE}
XtX = t(X) %*% X
print("XtX:")
print(XtX)
XXt = X %*% t(X)
print("XXt:")
print(XXt)
```


### Part (b)

```{r 5b, message=FALSE, warning=FALSE}
fit_XtX = eigen(XtX)
print(fit_XtX, digits=2)

fit_XXt = eigen(XXt)
print(fit_XXt, digits=2)
```

### Part (c)
```{r 5c, message=FALSE, warning=FALSE}
a = diag(fit_XtX$values, nrow=3, ncol=3)
b = matrix(fit_XtX$vectors, nrow = 3)

print(b %*% a %*% t(b))
```

$$ => X'X = \Gamma\Lambda\Gamma'$$

### Part (d)
```{r 5d, message=FALSE, warning=FALSE}

```

### Part (e)
```{r 5e, message=FALSE, warning=FALSE}

```

### Part (f)
```{r 5f, message=FALSE, warning=FALSE}

```

### Part (g)
```{r 5g}

```

### Part (h)
```{r 5h}

```

### Part (i)
```{r 5i}

```


# Problem 4

```{r}
R = matrix(c(1, .4919, .2636, .4653, -.2277, .0652, 
             .4919, 1, .3127, .3506, -.1917, .2045, 
             .2635, .3127, 1, .4108, .0647, .2493, 
             .4653, .3506, .4108, 1, -.2249, .2293, 
             -.2277, -.1917, .0647, -.2249, 1, -.2144,
             .0652, .2045, .2493, .2293, -.2144, 1), byrow=T, ncol=6,
           dimnames = list(c("Bluegill", "BlackCrappie", "SmallmouthBass",
                             "LargemouthBass","Walleye","NorthernPike"),
                           c("Bluegill", "BlackCrappie", "SmallmouthBass",
                             "LargemouthBass","Walleye","NorthernPike")))
print("Coorelation Matrix:")
print(R)
```

### Part (a)

```{r 4a, message=FALSE, warning=FALSE}
print("Correlation within the centrarchid family:")
print(R[1:4,1:4])
```

There seems to be fair amount of positive correlation between the centrarchid family of fish. Hence they seem to be living around each other.

Walleye on the other hand seems to have small negative correlation with the other fish, implying it must be residing far from others.


### Part (b)

```{r 4b, message=FALSE, warning=FALSE}
fit <- prcomp(R[1:4,1:4], scale=F)
plot(fit)

#First two loading vectors
print(fit$x[,1:2])
```

PC1 - distinguishes between Bass and other fish in the centrarchid family

PC2 - 
  * within the Bass family, distinguishes between large and small bass
  * within the non-bass family distinguishes between Bluegill and BlackCrappie


### Part (c)
```{r 4c, message=FALSE, warning=FALSE}
fit <- prcomp(R)
print("Scree Plot:")
plot(fit)
```


### Part (d)
```{r 4d, message=FALSE, warning=FALSE}
summary(fit)
```

The first two Principal Components explain 77.8% of the variation

### Part (e)
```{r 4e, message=FALSE, warning=FALSE}
print(fit$x[,1])
```

The first principal component isolates Walleye from the rest